---
title: "Linear models"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

### Problem 1

Loading data

```{r}

homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
  
```

Start with one city.

```{r}
baltimore_df=
  homicide_df %>% 
  filter(city_state=="Baltimore, MD")

glm(resolution~victim_age+ victim_race + victim_sex, 
    data=baltimore_df,
    family=binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR= exp(estimate),
    CI_lower=exp(estimate-1.96 * std.error),
    CI_upper=exp(estimate+1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits=3)
```

Try across cities.

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
  
```

Male vs Female Homicide Resolutions

```{r}
models_results_df %>% 
  filter(term=="victim_sexMale") %>% 
mutate(city_state=fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x=city_state, y=OR))+
  geom_point()+
  geom_errorbar(aes(ymin=CI_lower, ymax=CI_upper))+
  theme(axis.text.x= element_text (angle=90, vjust=0.5, hjust=1))
```


### Problem 2

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}

birth_weight_df = 
  read_csv("./data/birthweight.csv") %>%   
janitor::clean_names() %>% 
  mutate(
   frace = as.factor(frace),
   frace = recode(frace,
    "1" = "White", 
    "2" = "Black", 
    "3" = "Asian", 
    "4" = "Puerto Rican", 
    "8" = "Other", 
    "9" = "Unknown"),
   mrace = as.factor(mrace),
   mrace = recode(mrace, 
    "1" = "White", 
    "2" = "Black", 
    "3" = "Asian", 
    "4" = "Puerto Rican", 
    "8" = "Other"),
   babysex = as.factor(babysex),
   babysex = recode(babysex, 
     "1" = "male", 
    "2" = "female"),
   malform = as.factor(malform),
   malform = recode(malform,
    "0" = "absent", 
    "1" = "present")) %>% 
  filter(
    frace %in% c("White", "Black", "Puerto Rican")
  )
```


I propose that `blength`, `delwt`, `bhead`,`parity`, `mrace`,  `fincome`, `smoken`, and `pnumlbw` would be in the model. I chose baby's length at birth, baby's head circumference, mother's weight at delivery, mother's number of live births before this pregnancy,average number of cigarettes smoked per day during pregnancy,and previous number of low birth weight babies since these variables are relevant to gestational and newborn development. I chose race and family monthly income since they are socioeconomic factors that could impact baby's birthweight.

```{r}
birth_weight_df %>% 
  mutate(
    mrace=fct_infreq(mrace)
  )
```

Baby's Length

```{r}
length_fit = lm(bwt ~ blength, data = birth_weight_df)
length_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```


Baby's Head Circumference

```{r}
head_fit = lm(bwt ~ bhead, data = birth_weight_df)
head_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```


Mother's Weight at Delivery

```{r}
dlw_fit = lm(bwt ~ delwt, data = birth_weight_df)
dlw_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```


Mother's Number of Live Births Before Pregnancy

```{r}
parity_fit = lm(bwt ~ parity, data = birth_weight_df)
parity_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```


Average Number of Cigarettes Smoked per Day 

```{r}
smoke_fit = lm(bwt ~ smoken, data = birth_weight_df)
smoke_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```


Previous Number of Low Births Before Pregnancy

```{r}
pnumlbw_fit = lm(bwt ~ pnumlbw, data = birth_weight_df)
head_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```

Out of all of these linear models, the only non-significant one was Mother's Number of Live Births Before Pregnancy so I made a new model with the significant variables so far.

```{r}
together_fit=lm(bwt ~ bhead+ blength+ delwt + smoken, data=birth_weight_df) 

together_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```


Now I added in the sociodemographic factors to see if they also were related.

Mother Race

```{r}
mrace_fit = lm(bwt ~ mrace, data = birth_weight_df)
mrace_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```

Family Income
```{r}
fincome_fit = lm(bwt ~ fincome, data = birth_weight_df)
fincome_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```

Both family income and race were also significant predictors in the dataset so I added them to my final linear model. 

```{r}
final_fit=lm(bwt ~ bhead+ blength+ delwt + smoken+ mrace+fincome, data=birth_weight_df) 

final_fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits=3)
```

Plot of model residuals against fitted values

```{r}
residuals_df = 
  birth_weight_df %>% 
  modelr::add_residuals(final_fit) %>% 
  modelr::add_predictions(final_fit) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    x = "Predicted birthweight", 
    y = "Residuals", 
    title = "Residuals Against Fitted Values")

residuals_df
```
```

Compare your model to two others:


One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

### Problem 3

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

